{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-26 thesis/education.tex
change location of LFTs from 410user to landslide-friendly-tests
handle some todos, dont handle others

\noindent The tests can all be viewed at \url{https://github.com/bblum/landslide/tree/master/pebsim/p2-basecode/410user/progs}.

@@

% TODO check figure placement
\begin{table}[h]
	\begin{center}
		\small
		% note, smth like fisher's method which "combines independent studies into a meta p value"
		% cannot be applied for 2 reasons - 1, these are not independent since the same studence were
		% running landslide on diff test cases; 2, the hypotheses tested are different for each test
	\begin{tabular}{r||cc|cc|c|cc}
		& \multicolumn{2}{c|}{\bf Landslide users} & \multicolumn{2}{c|}{\bf Non-users} & & \multicolumn{2}{c}{\bf Significance} \\
		\bf Bug name & \bf \#correct & \bf \#buggy & \bf \#correct & \bf \#buggy & \bf $\Delta$\%correct & $p$ & \bf cutoff \\
		\hline
		% calculated in R with (rest done similarly)
		% fisher.test(rbind(c(11,10),c(0,16)))$p.value --> 0.0006143355
		Exit UAF  & 11	& 10	& 0	& 16	& $+52.4$\%	& 0.00061	& 0.0125 \\
		% I spent more hours than I should have looking at this landslide-user-submitted-bug group.
		% Can't name the group here, but if you're me and have the analysis file,
		% it's the one from f17 with 3 mutex_test logs in their "u" snapshot.
		% It seems like a mysterious older version of mutex.c was present between their "a" version
		% (which gets fully verified very quickly btw) and their "u" version (for which landslide finds
		% the bug in just a few minutes, no way they could be running the same version; plus the reported
		% data races and so on in the id-log don't line up to the "u" mutex.c source code).
		% That version was correct, then they updated it (introducing the bug) and ran a bunch of other tests,
		% but didn't rerun mutex_test to check their update.
		Mutex     & 17	& 1	& 19	& 0	& $-05.6$\%	& 0.49 & 0.0125 \\
		Paraguay  & 20	& 1	& 14	& 2	& $+07.7$\%	& 0.57	& 0.0125 \\
		% nb this is any downgrade related ding not just dgr tested ones
		% nb 1 is msising bc didn't impl rwlock at all
		% nice p value
		Downgrade & 10	& 3	& 19	& 4	& $-05.7$\%	& 0.69	& 0.0125 \\ % TODO verify these buggers by eye
		\hline
		\bf Total & 58	& 15	& 52	& 22	& $+09.2$\%	& 0.26 & 0.05 \\
	\end{tabular}
	\end{center}
	\caption[Correlation of student Landslide use with solving certain bugs.]
		{
		Correlation of student Landslide use with solving certain bugs in their final submission during Fall 2017 and Spring 2018 semesters.
		%$\Delta P$ stands for the increased probability that a student submits a correct P2
		%(with respect to each bug)
		%given that they use Landslide, versus not using Landslide; i.e.,
		%$P(\mathsf{correct}|\mathsf{landslide})-
		%P(\mathsf{correct}|\mathsf{non\text{-}user})$.
		Note that the totals in the bottom row double-count students,
		which makes sense only if you believe the incidences of each bug in a given submission are independent.
		}
	\label{tab:eval-common-bugs}
\end{table}

@@

Further investigating the one group who submitted a buggy mutex,
I found that while they had run {\tt mutex\_test} in Landslide
(and even found and fixed a separate deterministic bug already),
Landslide found no bug in what was presumably a correct implementation,
then they updated % oops
their code, introducing the bug, without testing it again thereafter.
% TODO: some reference to future work about focused testing on small code diffs? talked with some industry ppl about this sometimes

>>>>

\noindent The tests can all be viewed at
%\url{https://github.com/bblum/landslide/tree/master/pebsim/p2-basecode/410user/progs}.
\url{https://github.com/bblum/landslide/tree/master/pebsim/p2-basecode/landslide-friendly-tests}.

@@

% TODO check figure placement
\begin{table}[h]
	\begin{center}
		\small
		% note, smth like fisher's method which "combines independent studies into a meta p value"
		% cannot be applied for 2 reasons - 1, these are not independent since the same studence were
		% running landslide on diff test cases; 2, the hypotheses tested are different for each test
	\begin{tabular}{r||cc|cc|c|cc}
		& \multicolumn{2}{c|}{\bf Landslide users} & \multicolumn{2}{c|}{\bf Non-users} & & \multicolumn{2}{c}{\bf Significance} \\
		\bf Bug name & \bf \#correct & \bf \#buggy & \bf \#correct & \bf \#buggy & \bf $\Delta$\%correct & $p$ & \bf cutoff \\
		\hline
		% calculated in R with (rest done similarly)
		% fisher.test(rbind(c(11,10),c(0,16)))$p.value --> 0.0006143355
		Exit UAF  & 11	& 10	& 0	& 16	& $+52.4$\%	& 0.00061	& 0.0125 \\
		% I spent more hours than I should have looking at this landslide-user-submitted-bug group.
		% Can't name the group here, but if you're me and have the analysis file,
		% it's the one from f17 with 3 mutex_test logs in their "u" snapshot.
		% It seems like a mysterious older version of mutex.c was present between their "a" version
		% (which gets fully verified very quickly btw) and their "u" version (for which landslide finds
		% the bug in just a few minutes, no way they could be running the same version; plus the reported
		% data races and so on in the id-log don't line up to the "u" mutex.c source code).
		% That version was correct, then they updated it (introducing the bug) and ran a bunch of other tests,
		% but didn't rerun mutex_test to check their update.
		Mutex     & 17	& 1	& 19	& 0	& $-05.6$\%	& 0.49 & 0.0125 \\
		Paraguay  & 20	& 1	& 14	& 2	& $+07.7$\%	& 0.57	& 0.0125 \\
		% nb this is any downgrade related ding not just dgr tested ones
		% nb 1 is msising bc didn't impl rwlock at all
		% nice p value
		Downgrade & 10	& 3	& 19	& 4	& $-05.7$\%	& 0.69	& 0.0125 \\
		\hline
		\bf Total & 58	& 15	& 52	& 22	& $+09.2$\%	& 0.26 & 0.05 \\
	\end{tabular}
	\end{center}
	\caption[Correlation of student Landslide use with solving certain bugs.]
		{
		Correlation of student Landslide use with solving certain bugs in their final submission during Fall 2017 and Spring 2018 semesters.
		%$\Delta P$ stands for the increased probability that a student submits a correct P2
		%(with respect to each bug)
		%given that they use Landslide, versus not using Landslide; i.e.,
		%$P(\mathsf{correct}|\mathsf{landslide})-
		%P(\mathsf{correct}|\mathsf{non\text{-}user})$.
		Note that the totals in the bottom row double-count students,
		which makes sense only if you believe the incidences of each bug in a given submission are independent.
		}
	\label{tab:eval-common-bugs}
\end{table}

@@

Further investigating the one group who submitted a buggy mutex,
I found that while they had run {\tt mutex\_test} in Landslide
(and even found and fixed a separate deterministic bug already),
Landslide found no bug in what was presumably a correct implementation,
then they updated % oops
their code, introducing the bug, without testing it again thereafter.
% not-to-do: some reference to future work about focused testing on small code diffs? talked with some industry ppl about this sometimes

### 2018-12-26 thesis/tm.tex
handle some todos, dont handle others

\revision{If the user supplies the {\tt -W} option (\cref{sec:tm-interface}),
Landslide will emulate STM's weak atomicity semantics (\cref{sec:tm-weak-atomicity}).
%to allow non-transactional code
%to interleave between a {\tt \_xbegin()}/{\tt \_xend()} pair.
Concretely, it achieves this by
reporting both halves of data races that occur between transactional and non-transactional code (\cref{sec:landslide-datarace})
(previously, only the latter half would be reported),
allowing preemption points to be identified during transactions (\cref{sec:landslide-pps}),
and allowing system calls during transactions to proceed rather than forcing a {\tt CAPACITY} abort
(as programmers may now need to nest other synchronization within transactions to protect against non-transactional code).
%and traversing the interleaving arbitrarily far back from a {\tt \_xend()} to find the appropriate preemption
%(previously, the xbegin would be at the immediately previous PP)
Landslide requires the other STM options ({\tt -A -S}) also to be enabled to emulate weak atomicity,
partly because I know of no existing HTM system which implements it,
but mostly to keep the implementation simple.
As one might expect, it inflates state spaces considerably, % TODO: measure it with htm2 or some such?
and moreover,
every single HTM benchmark in the upcoming evaluation,
implemented to rely on strong atomicity,
exhibits bugs under weak atomicity.
Commit 68401fc implements this mode.}

>>>>

\revision{If the user supplies the {\tt -W} option (\cref{sec:tm-interface}),
Landslide will emulate STM's weak atomicity semantics (\cref{sec:tm-weak-atomicity}).
%to allow non-transactional code
%to interleave between a {\tt \_xbegin()}/{\tt \_xend()} pair.
Concretely, it achieves this by
reporting both halves of data races that occur between transactional and non-transactional code (\cref{sec:landslide-datarace})
(previously, only the latter half would be reported),
allowing preemption points to be identified during transactions (\cref{sec:landslide-pps}),
and allowing system calls during transactions to proceed rather than forcing a {\tt CAPACITY} abort
(as programmers may now need to nest other synchronization within transactions to protect against non-transactional code).
%and traversing the interleaving arbitrarily far back from a {\tt \_xend()} to find the appropriate preemption
%(previously, the xbegin would be at the immediately previous PP)
Landslide requires the other STM options ({\tt -A -S}) also to be enabled to emulate weak atomicity,
partly because I know of no existing HTM system which implements it,
but mostly to keep the implementation simple.
As one might expect, it inflates state spaces considerably,
and moreover,
every single HTM benchmark in the upcoming evaluation,
implemented to rely on strong atomicity,
exhibits bugs under weak atomicity.
Commit 68401fc implements this mode.}
